---
title: "Correlating Like a Data Master"
author: "Anjani Sowmya Bollapragada - 23851219"
format: html
editor: visual
---

### Purpose of this chapter

Assess relationships within a novel data set using publication quality tables and plots

### Required SetUp

```{r}
if (!require(pacman))
  install.packages("pacman")

pacman::p_load(colorblindr,
       dlookr,
       formattable,
       GGally,
       ggdist,
       ggpubr,
       ggridges,
       here,
       tidyverse)

# Set global ggplot() theme
# Theme pub_clean() from the ggpubr package with base text size = 16
theme_set(theme_pubclean(base_size = 12)) 
# All axes titles to their respective far right sides
theme_update(axis.title = element_text(hjust = 1))
# Remove axes ticks
theme_update(axis.ticks = element_blank()) 
# Remove legend key
theme_update(legend.key = element_blank())
```

### Load an Examine a Data Set

```{r}
# Let's load the canopy tree data set
dataset <- read.csv("week19_airline_safety.csv")

# What does the data look like?
dataset |>
  head() |>
  formattable()
```

### Describe and Visualize Correlations

Correlations are a statistical relationship between two numerical variables, may or may not be causal. Exploring correlations in your data allows you determine data independence, a major assumption of parametric statistics, which means your variables are both randomly collected.

```{r}
# Table of correlations between numerical variables (we are sticking to the default Pearson's r coefficient)
correlate(dataset) |>
  formattable()
```

```{r}
# Correlation matrix of numerical variables
dataset |>
plot_correlate()
```

### Visualize Correlations within Groups

```{r}
dataset |>
  group_by(airline) |>
  plot_correlate()
```

### Describe and Visualize Relationships Based on Target Variables

#### Target Variables

*Target variables* are essentially numerical or categorical variables that you want to relate others to in a data frame. *dlookr* does this through the *target_by()* function, which is similar to *group_by()* in dplyr. The *relate()* function then briefly analyzes the relationship between the target variable and the variables of interest.

#### Numerical Target Variables: Numerical Variable of Interest

```{r}
# First, we need to remove NAs, they cause an error
dataset.noNA <- dataset |> 
  drop_na()

# The numerical predictor variable that we want
num <- target_by(dataset.noNA, n_events)

# Relating the variable of interest to the numerical target variable
num_num <- relate(num, avail_seat_km_per_week)

# Summary of the regression analysis - the same as the summary from lm(Formula)
summary(num_num)
```

```{r}
# Plotting the linear relationship
plot(num_num)
```

#### Numerical Target Variables: Categorical Variable of Interest

```{r}
# The categorical predictor variable that we want
num <- target_by(dataset, n_events) 

# We need to change Group to a factor
num$airline <- as.factor(num$airline)

# Relating the variable of interest to the numerical target variable
num_cat <- relate(num, airline)

# Summary of the ANOVA analysis - the same as the summary from anova(lm(Formula))
summary(num_cat)
```

```{r}
plot(num_cat) + 
  theme(axis.text.x = element_blank())
```


#### Categorical Target Variables: Numerical Variable of Interest

```{r}
# The categorical predictor variable that we want
categ <- target_by(dataset, airline)

# Relating the variable of interest to the numerical target variable
cat_num <- relate(categ, n_events)

# Summary of descriptive statistics
summary(cat_num)
```

```{r}
plot(cat_num) 
```

#### Categorical Target Variables: Categorical Variable of Interest

```{r}
# Create new categorical column
cat_dataset <- dataset |>
  select(n_events, airline) |>
  drop_na() |>
  mutate(High_events = ifelse(
    n_events > (mean(n_events + sd(n_events))), 
                          "Yes", 
                          "No"))

# New dataset 
cat_dataset |>
  head() |>
  formattable()
```

```{r}
# The categorical predictor variable that we want
categ <- target_by(cat_dataset, High_events)

# Relating the variable of interest to the categorical target variable
cat_cat <- relate(categ, airline)

# Summary of the Chi-square test for Independence
summary(cat_cat)
```

```{r}
plot(cat_cat)
```

